\huge \lr: \\ \centerline{an interactive assistant for \ma}

### 2023-09-27 writing/AMCTM2023/presentation.tex
[ AMCTM23 ] minor fixes

\begin{frame}{Code generation}
There are lot of menial programming tasks in the wild\pause
\begin{itemize}[<+->]
  \item e.g., code that handles data serialisation
  \begin{itemize}
    \item have to keep in mind what input is required, what are the dimensions, what format, etc...
    \item but this is meta information, not necessarily evident in the \ma{} implementation
  \end{itemize}
\end{itemize}
\visible<5->{We can do better}
\pause
\begin{itemize}[<+->]
  \item develop a high-level \keyword{metalanguage of descriptions}
  \item read and validate descriptions, make sanity checks and generate \ma{} code that closely resembles hand crafted one
\end{itemize}
\end{frame}

@@

\begin{frame}{Elaborating Matlab expressions to CoreTT}
  \begin{columns}
  \begin{column}{0.3\textwidth}
    \begin{tikzpicture}[
      start chain=going below,
      every join/.style={fadearrow},
      node distance=0.6cm
      ]
      \node (s0) [fade, on chain,join] {Parsing};
      \node (s1) [highlight, on chain, join] {Elaboration};
      \node (s2) [fade, on chain, join] {Typechecking};
      \node (s3) [fadeend, on chain, join] {Other stuff};
    \end{tikzpicture}
    \begin{tikzpicture}[remember picture, overlay]
      \node [xshift=0.12\paperheight,yshift=0.15\paperheight] at (current page.south west){
        \includegraphics[width=0.2\paperwidth, keepaspectratio]{mate.png}
      };
    \end{tikzpicture}
  \end{column}
  \begin{column}{0.55\textwidth}
    The \keyword{interface} between the type system and the user written code. \pause
    \begin{itemize}[<+->]
      \item tries to fill the gaps between the two
      \begin{itemize}
        \item e.g., comes up with types for \ma{} expressions
      \end{itemize}
      \item makes sure the type errors are human readable
    \end{itemize}
  \end{column}
\end{columns}
\end{frame}

\begin{frame}[fragile]{Elaborating Assignments}
  \begin{lstlisting}[xleftmargin=10em]
    %> n!\tikzmark{ndecl}! :: int!\tikzmark{intdecl}!
    n !\tikzmark{nlhs}!=!\tikzmark{ass1}! 5!\tikzmark{five}!;
    x !\tikzmark{xlhs}!=!\tikzmark{ass2}! n !\tikzmark{nrhs}!
    %> typeof x!\tikzmark{xtypeof}!
  \end{lstlisting}
  \pause
  \mycallout<+>{ndecl}{see a new variable $n$}
  \mycallout<+>{intdecl}{simplify $n$'s type}
  \mycallout<+>{ndecl}{declare $n$ in scope with type $int$}
  \mycallout<+>{ass1}{see an assignment at yet unknown type}
  \mycallout<+>{five}{process the RHS, $5$, using the the type of LHS $n$}
  \mycallout<+>{nlhs}{find the type of the declared $n$}
  \mycallout<+>{ass1}{constrain the type of RHS,$5$, to be equal to that of LHS, $n$}
  \mycallout<+>{ass2}{see an assignment at yet unknown type}
  \mycallout<+>{nrhs}{simplify the RHS, $n$, finding its type}
  \mycallout<+>{xlhs}{declare $x$ in scope with unknown type}
  \mycallout<+>{ass2}{constrain the type of RHS, $n$, to be equal to that of LHS, $x$}
  \mycallout<+>{xtypeof}{find the type of the declared $x$}
\end{frame}

\begin{frame}[fragile]{Current Progress}
  \begin{itemize}[<+->]
    \item The code is available at \repo
    \item We are currently parsing a lot of common \ma{} constructions
    \item Directives for renaming and code generation are ready
    \item Implemented most of the underlying type system for \ma
    \item Actively working on elaboration
  \end{itemize}
  \pause
  \vspace{3em}
  \LARGE \centerline{Thank you for your attention!}
\end{frame}

>>>>

\begin{frame}{Code generation}
There are lot of menial programming tasks in the wild\pause
\begin{itemize}[<+->]
  \item e.g., code that handles data serialisation
  \begin{itemize}
    \item have to keep in mind what input is required, what are the dimensions, what format, etc...
    \item but this is metainformation, not necessarily evident in the \ma{} implementation
  \end{itemize}
\end{itemize}
\visible<5->{We can do better}
\pause
\begin{itemize}[<+->]
  \item develop a high-level \keyword{metalanguage of descriptions}
  \item read and validate descriptions, make sanity checks and generate \ma{} code that closely resembles hand crafted one
\end{itemize}
\end{frame}

@@

\begin{frame}{Elaborating Matlab expressions to CoreTT}
  \begin{columns}
  \begin{column}{0.3\textwidth}
    \begin{tikzpicture}[
      start chain=going below,
      every join/.style={fadearrow},
      node distance=0.6cm
      ]
      \node (s0) [fade, on chain,join] {Parsing};
      \node (s1) [highlight, on chain, join] {Elaboration};
      \node (s2) [fade, on chain, join] {Typechecking};
      \node (s3) [fadeend, on chain, join] {Other stuff};
    \end{tikzpicture}
    \begin{tikzpicture}[remember picture, overlay]
      \node [xshift=0.12\paperheight,yshift=0.15\paperheight] at (current page.south west){
        \includegraphics[width=0.2\paperwidth, keepaspectratio]{mate.png}
      };
    \end{tikzpicture}
  \end{column}
  \begin{column}{0.55\textwidth}
    The \keyword{interface} between the type system and the user written code. \pause
    \begin{itemize}[<+->]
      \item tries to fill the gaps between the two
      \begin{itemize}
        \item e.g., comes up with types for \ma{} expressions
      \end{itemize}
      \item makes sure that type errors are human readable
    \end{itemize}
  \end{column}
\end{columns}
\end{frame}

\begin{frame}[fragile]{Elaborating Assignments}
  \begin{lstlisting}[xleftmargin=10em]
    %> n!\tikzmark{ndecl}! :: int!\tikzmark{intdecl}!
    n !\tikzmark{nlhs}!=!\tikzmark{ass1}! 5!\tikzmark{five}!;
    x !\tikzmark{xlhs}!=!\tikzmark{ass2}! n !\tikzmark{nrhs}!
    %> typeof x!\tikzmark{xtypeof}!
  \end{lstlisting}
  \pause
  \mycallout<+>{ndecl}{see a new variable $n$}
  \mycallout<+>{intdecl}{simplify $n$'s type}
  \mycallout<+>{ndecl}{declare $n$ in scope with type $int$}
  \mycallout<+>{ass1}{see an assignment at yet unknown type}
  \mycallout<+>{five}{process the RHS, $5$, using the the type of LHS $n$}
  \mycallout<+>{nlhs}{find the type of the declared $n$}
  \mycallout<+>{ass1}{constrain the type of RHS, $5$, to be equal to that of LHS, $n$}
  \mycallout<+>{ass2}{see an assignment at yet unknown type}
  \mycallout<+>{nrhs}{simplify the RHS, $n$, finding its type}
  \mycallout<+>{xlhs}{declare $x$ in scope with unknown type}
  \mycallout<+>{ass2}{constrain the type of RHS, $n$, to be equal to that of LHS, $x$}
  \mycallout<+>{xtypeof}{find the type of the declared $x$}
\end{frame}

\begin{frame}[fragile]{Current Progress}
  \begin{itemize}[<+->]
    \item The code is available at \repo
    \item We are currently parsing a lot of common \ma{} constructions
    \item Directives for renaming and code generation are working
    \item Implemented most of the underlying type system for \ma
    \item Actively working on elaboration
  \end{itemize}
  \pause
  \vspace{3em}
  \LARGE \centerline{Thank you for your attention!}
\end{frame}
